---
title: "Algunos elementos visuales sencillos y lucidores"
description: "Análisis de los datos más bonito"
date: "2026-09-15"
---

::: {.callout-note}
Note that there are five types of callouts, including:
`note`, `warning`, `important`, `tip`, and `caution`.
:::

::: {.callout-tip}
## Tip with Title

This is an example of a callout with a title.
:::

::: {.callout-caution collapse="true"}
## Expand To Learn About Collapse

This is an example of a 'folded' caution callout that can be expanded by the user. You can use `collapse="true"` to collapse it by default or `collapse="false"` to make a collapsible callout that is expanded by default.
:::

::: {.callout-tip title="Tip with Title"}
This is a callout with a title.
:::

::: {.callout-note appearance="simple"}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention.

:::

::: {.callout-note icon=false}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention.

:::

```{r}
library(ggplot2)
library(ggalluvial)
```

+ La librería `{ggalluvial}` nos ayuda a mostrar cómo fluyen categorías entre etapas

```{r}
datos <- data.frame(
  carrera_inicial = c("Ing.", "Ing.", "Diseño", "Diseño", "Negocios", "Negocios"),
  carrera_final   = c("Ing.", "Datos", "Diseño", "Datos", "Negocios", "Datos"),
  n = c(40, 15, 30, 10, 25, 20)
)
```

```{r}
datos |> head()
```

Hagamos una grafiquita sencilla


```{r}
ggplot(datos,
       aes(axis1 = carrera_inicial, axis2 = carrera_final, y = n)) +
  geom_alluvium(aes(fill = carrera_inicial)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Año 1", "Año 2")) +
  labs(title = "Cambio de carrera entre año 1 y año 2", y = "Estudiantes") +
  theme_minimal()
```
