---
title: "Algunos elementos visuales sencillos y lucidores"
description: "Análisis de los datos más bonito"
date: "2026-09-15"
---

::: {.callout-note}
Note that there are five types of callouts, including:
`note`, `warning`, `important`, `tip`, and `caution`.
:::

::: {.callout-tip}
## Tip with Title

This is an example of a callout with a title.
:::

::: {.callout-caution collapse="true"}
## Expand To Learn About Collapse

This is an example of a 'folded' caution callout that can be expanded by the user. You can use `collapse="true"` to collapse it by default or `collapse="false"` to make a collapsible callout that is expanded by default.
:::

::: {.callout-tip title="Tip with Title"}
This is a callout with a title.
:::

::: {.callout-note appearance="simple"}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention.

:::

::: {.callout-note icon=false}

## Pay Attention

Using callouts is an effective way to highlight content that your reader give special consideration or attention.

:::

```{r}
library(ggplot2)
library(ggbump)
```

Con la librería `{ggbump}` se hacen "bump charts" que muestran cómo cambian rankings a lo largo del tiempo (típico para ranking de equipos, ventas por región, etc.).

Nuestros datos:

```{r}
datos <- data.frame(
  pais = rep(c("México", "Brasil", "Argentina", "Chile"), each = 3),
  anio = rep(c(2022, 2023, 2024), times = 4),
  ranking = c(1, 2, 1,   # México
              2, 1, 3,   # Brasil
              3, 4, 2,   # Argentina
              4, 3, 4)   # Chile
)
```

```{r}
datos |> head(n = 10)
```

Hagamos una grafiquita sencilla

```{r}
ggplot(datos, aes(x = anio, y = ranking, color = pais)) +
  geom_bump(size = 1.5, smooth = 8) +
  geom_point(size = 4) +
  scale_y_reverse(breaks = 1:4) +
  labs(title = "Cambio de ranking por año",
       x = "Año", y = "Posición") +
  theme_minimal()
```
